---
title: "DSA Individual Assignment by Saurabh Gupta (200101134)"
author: 'Submited to: Prof.(Dr.) Sridhar Vaithianathan, Associate Professor (Analytics).IMT Hyderabad'
date: "19/12/2020"
output: word_document
---
# Table of Content 
# 1) Mathematical Operations & Variable in R
# 2) Data Types in R
# 3) Function in R
# 4) Data Structures in R which are as follows
# a) Vector
# b) Dataframe
# c) Matrix
# d) Array 
# e) List
# 5) Reading Files in R and Built in dataset
# 6) Descriptive Statistics
# 7) T-test & Annova
Mathematical Operations & Variable in R

```{r}
# Commenting (# is used to comment i.e. sentences after 
# will not be included in the code)

# Ctrl + L is used to delete all the contents of the console
# Ctrl/command + Enter - To execute the program select the line and 
# and press Ctrl/command +Enter.

# Creating/assinging a variable:
x = 5 #(The variable x is assigned the value of 5)
y = 7 #(The variable y is assigned the value of 7)

# Mathematical Operations in Python
x+y #addition of two vaiable
x-y #subtration of two variable
x*y #multiplication of two variable
x/y # Division of two variable
sqrt(x) # Sqareroot of x
x^y #x to the power of y
exp(x) #exponential value of x
log(x, base = exp(1)) # Log value of x when base is 1
log10(x) # Log value when the base is 10
factorial(x) # factorial value of x
cos (x) #Cos value of x
abs(x) #Absoulute value i.e. without any sign value of x
```

Data Types in R
```{r}
# There are 4 data types in R

# 1) Numeric - Numbers which can be further divided into
#integer and whole numbers

a = 1 
class(a) # class is used to find the class i.e. the data 
#type of the variable

b = 1L #L is used to specify the variable as Integer numberic
class(b) # to find the data type of vaiable b

# 2) Character - Words/strings (Nominal) & Classification (ordinal)
c = "Saurabh" # "" is used to specify as Character
class(c) # to find the data type of variable c
# class must be written in lower case so as to work as a function

# 3) Logical - True (1) & False (0)
TRUE * 3 # Both True & False should be written in Capitals letters
FALSE * 3 # Both True & False should be written in Capitals letters

# 4) Date
date1 = as.Date("2020-12-20")
date1
class(date1)
as.numeric(date1) #Vonverts date into numeric number

# Date and Time 

date2 = as.POSIXct("2020-12-20 12:00")
date2
class(date2)
as.numeric(date2)
```

Function in R
```{r}
getwd() # to get the working directory i.e. the place the files are being stored

# Functions in R
divider = function(x,y) { #divider is a variable name
  result = x/y
  print(result) # Print is used to get the result of the code
}
divider(50,25)
divider (100,25)

# Multiplication
multiply = function(a,b){ # Multiply is a vaiable name
  result = a * b
  print (result)
}
multiply(23,25)
multiply (19,20)


# Variables Names are CASE SENSITIVE
X=10
x=24

# CONCATENATION AND ARRAYS
f <- c(1,2,3,4,5) 
f = c(1,2,3,4,5)
f
f+4

d = f / 4
d

f+d

f = c(1,2,3,4,5)

# Listing and Deleting Objects (Variables)
ls()
rm (a)
rm (list = ls())
```

Data Structures in R
```{r}
# There are 5 types of data structure in R
# Data Structures in R
# Data come in many types and structures which can pose a problem for some...
# ...analysis environments but R handles them with ease.

## VECTOR
# The most common data structure is the one-dimensional vector
# Vector forms the basis of everything in R.
# A vector is collection of elements of same type.
# (ie) A vector cannot be of mixed type.
# R is a Vectorized Language. That means operations are applied to each element of the vector automatically,
# .., without the need to loop through the vector.
# This is a powerful concept and vector plays a crucial and significant role in R.
# Refer the file : "3 All about Vectors in R - 26 Nov" (R File)

## DATA FRAME
# Data Frames(DF) - Most useful features of R & also cited reason for R's ease of use.
# In dataframe, each column is actually a vector, each of which has same length.
# Each column can hold different type of data.
# Also within each column, each element must be of same type, like vectors.
# Refer the file : "4 DSA Data Structures_Data.Frame - 4 Dec" (R File)

## MATRICES
# A matrix (plural matrices) is a rectangular array or table of numbers, symbols, or expressions...
#..., arranged in rows and columns.(i.e.) 2-Dimensional Array
# Similar to data.frame(RxC) and also similar to Vector
# Matrix - Element by element operations are possible.
# Refer the file : "4 DSA Data Structures_Matrices - 4 Dec" (R File)

## ARRAYS
# Arrays - An array is essentially a multidimensional vector.
# It must all be of the same type and 
# ...individual elements are accessed using Square Brackets.
# First element is Row(R) Index, Second Element is Column(C) Index and 
# the remaining elements are for Outer Dimensions (OD).
# Refer the file : "4 DSA Data Structures_Arrays - 4 Dec" (R File)

## LIST
# Lists - Stores any number of items of any type.
# List can contain all numerics or characters or...
#...a mix of the two or data.frames or recursively other lists.
# Refer the file : "4 DSA Data Structures_List - 4 Dec" (R File)
```

1) Vector
```{r}
# Vector - R is called as Vectorized language.
v = c(1,2,3,4,5)
s = v*2
s
# A vector is collec tion of elements, all of same type.
# A vector canot be of mixed type.
# Vector Operation
d = v-2
d
f = v /2
f
sqrt(f)
# A vector is collection of elements of same type.
# (ie) A vector cannot be of mixed type.
# R is a Vectorized Language. Thant means operations are applied to each element of the vector automatically,
# .., without the need to loop through the vector.
# This is a powerful concept and vector plays a crucial and significant role in R.

# Creating Vectors
# The most common way to create a Vector is using 'c' [combine]
x = c(1,2,3,4,5,6,7,8,9,10)
x

# Vector Operations
x*3 # multiplies each element by 3; No loops necessary!
x+2
x-3
x/4
x^2
sqrt(x)

# colon (:)  operation - Sequencing
# Creates sequence of Numbers in either direction!
1:10 #(: - Through)
10:1
-2:3
5:-7

# More on Vector Operations ... Two vectors
# create two vectors of equal length
x = 1:10
y = -5:4
x + y # Add
x-y
x*y
x/y
x^y
# check the length of each vector
length(x)
length(y)
# Unequal length vectors
x+c(1,2) # Shorter vector gets recycled!
x+c (1,2,3)# If Longer vector is not "multiple" of shorter vector, a warning is given!
# Comparison also work on vector!
x <= 5
x<y
# Vector Comparison - "any" and "all"
x = 10:1
y = -4:5
any(x<y)
all(x<y)

# The "nchar" function also acts on each element of vector.
q = c("Hockey","Football","Baseball","Curlin","Rugby","Lacrosse",
      "Basketball","Tennis","Cricket","Soccer")
q
nchar(q)
nchar(y)

?nchar()

# Subscripting:Accessing "individual elements" in vector is done using square brackets []
x[1]
x[1:2]
x[c(1:5,9)]

# Give Names to Vector!
c(One = "a", Two = "y", Last = "r") # Name-Value pair
# You can Name the vector after creating vector as well!
w = 1:3
names(w) = c("a","b","c")
w
```

2) Dataframe
```{r}
#2) Dataframe - Combination of different vectors
# Sometimes data requires more complex storage than simple vectors.
# Data Structures - Apart from Vectors, we have Data Frames, Matrix, List and Array.

# Data Frames(DF) - Most useful features of R & also cited reason for R's ease of use.
# In dataframe, each column is actually a vector, each of which has same length.
# Each column can hold different type of data.
# Also within each column, each element must be of same type, like vectors.

# Creating a Dataframe from vectors

x = 10:1
y = -4:5
q = c("Hockey","Football","Baseball","Curlin","Rugby","Lacrosse",
      "Basketball","Tennis","Cricket","Soccer")
theDF = data.frame(x,y,q) # this would create a 10x3 data.frame with x, y and q as variable names
theDF

# Assigning Names
theDF = data.frame (First=x, Second =y, Sport = q)
theDF

# Checking the dimensions of the DF.
nrow(theDF)
ncol(theDF)
dim(theDF)
names (theDF)
names(theDF)[3]
rownames(theDF)

# Head and Tail
head(theDF)
head(theDF, n=7)
tail(theDF)

class(theDF)

# Accessing Individual Column using $
theDF$Sport # gives the third column named Sport

# Accessing Specific row and column
theDF[3,2] # 3rd row and 2nd Column
theDF[3,2:3] # 3rd Row and column 2 thru 3
theDF[c(3,5), 2]# Row 3&5 from Column 2; 
# since only one column was selected, it was returned as vector and hence no column names in output.

# Rows 3&5 and Columns 2 through 3
theDF[c(3,5), 2:3]

theDF[ ,3] # Access all Rows for column 3
theDF[ , 2:3]

theDF[2,]# Access all columns for Row 2
theDF[2:4,]
theDF[ , c("First", "Sport")]# access using Column Names
theDF[ ,"Sport"]# Access specific Column
class(theDF[ ,"Sport"])

theDF["Sport"]# This returns the one column data.frame
class(theDF["Sport"]) # Data.Frame

theDF[["Sport"]]#To access Specific column using Double Square Brackets
class(theDF[["Sport"]]) # Factor

theDF[ ,"Sport", drop = FALSE]# Use "Drop=FALSE" to get data.fame with single sqaure bracket.
class(theDF[ ,"Sport", drop = FALSE]) # data.frame

theDF[ ,3, drop = FALSE]
class(theDF[ ,3, drop = FALSE]) # data.frame

# To see how factor is stored in data.frame
newFactor = factor(c("Pennsylvania","New York","New Jersey","New York","Tennessee","Massachusetts","Pennsylvania","New York"))
newFactor
```

3) Matrix
```{r}
# Similar to data.frame(RxC) and also similar to Vector
# Matrix - Element by element operations are possible
V = matrix(1:10, nrow=5)# Create a 5x2 matrix
V
# A matrix (plural matrices) is a rectangular array or table of numbers, symbols, or expressions...
#..., arranged in rows and columns.(i.e.) 2-Dimensional Array

# Similar to data.frame(RxC) and also similar to Vector
# Matrix - Element by element operations are possible

A = matrix(1:10, nrow=5)# Create a 5x2 matrix
B = matrix(21:30, nrow=5)#Create another 5x2 matrix 
C = matrix (21:40, nrow=2)#Create another 2x10 matrix

A
B
C

nrow(A)
ncol(A)
dim(A)

# Add Them
A+B

# Multiply Them (Vector Multiplication!)
A
B
A*B  # A = 5x2 and B = 5x2

#See if the elements are equal
A == B

# Matrix Multiplication(MM. A is 5x2. B is 5x2. B-transpose is 2x5
A %*% t(B) #t is used for tranpose of a matrix

# Naming the Columns and Rows 
colnames(A)
rownames(A)
colnames(A)= c("Left","Right")
rownames(A)= c("1st","2nd","3rd","4th","5th")
colnames(B)
rownames(B)
colnames(B)= c("First","Second")
rownames(B)= c("One","Two","Three","Four","Five")
colnames(C)
rownames(C)

colnames(C) = LETTERS [1:10]
rownames(C) = c("Top", "Bottom")

# Matrix Multiplication. A is 5x2 and C is 2x10
dim(A) # finding the Rows(R)*C(Colomuns) of A 
dim(C) # Finding the R*C of C
t(A) # Transpose of A
A %*% C # Matrix Multiplication uses "%*%"
```

4) Arrays
```{r}
# 4) Arrays
# Arrays - An array is essentially a multidimensional vector.
# It must all be of the same type and 
# ...individual elements are accessed using Square Brackets.
# First element is Row(R) Index, Second Element is Column(C) Index and 
# the remaining elements are for Outer Dimensions (OD).

theArray = array(1:12, dim=c(2,3,2))# Total Elements = R x C x OD
theArray
theArray [1, ,]# Accessing all elements from Row 1, all columns, all outer dimensions & build C x OD (R x C)
theArray[1, ,1]# Accessing all elements from Row 1, all columns, first outer dimension 
theArray[, ,1]# Accessing all rows, all columns, first outer dimension
```

5) List
```{r}
# 5) List
# Lists - Stores any number of items of any type.
# List can contain all numerics or characters or...
#...a mix of the two or data.frames or recursively other lists.

# Lists are created with the "list" function.
# Each argument in "list" becomes an element of the list.

list(1,2,3)# creates a three element list
list(c(1,2,3))# creates a single element(vector with three elements)
list3 = list(c(1,2,3), 3:7)# create two element list
# first is three elements vector, next is five element vector.
list3
# The same can be written as
(list3 = list(c(1,2,3), 3:7))

# Two Element list
# First element is data.frame and next is 10 element vector
list(theDF, 1:10)# theDF is already created in previous exercise!

# Three element list
list5 = list(theDF, 1:10, list3)
list5

#Naming List (similar to column name in data.frame) 
names(list5)= c("data.frame", "vector","list")
names(list5)
list5
#Naming using "Name-Value" pair
list6 = list(TheDataFrame = theDF, TheVector = 1:10, TheList = list3)
names(list6)
list6

# Creating an empty list
(emptylist = vector(mode="list", length =4))

# Accessing individual element of a list - Double Square Brackets
# specify either element number or name
list5[[1]]
list5[["data.frame"]]
list5[[1]]$Sport
list5[[1]][,"Second"]
list5[[1]][,"Second", drop = FALSE]

# LENGTH OF LIST
length(list5)
names(list5)
list5
```

Reading Files in R and Built in dataset
```{r}
# Reading Files in R
# Its time that we load data in R.
# Most common way to get data is reading comma separated values(CSV)

# Reading CSVs
theUrl = "http://www.jaredlander.com/data/Tomato%20First.csv"
# visit https://www.jaredlander.com/data/ for other Datasets
tomato = read.table(file=theUrl, header=TRUE, sep =",")
head(tomato)

#It might be tempting to use read.csv but that is more trouble than it is worth,
#...and all it does is call read.table with some arguments preset.

# Sometimes CSVs(or tab delimited files) are poorly built, 
# where the cell separator has been used inside a cell.
# In this case read.csv2(or read.delim2)should be used instead of read.table.

# Reading Excel Data - Not worth the Effort.
# Unfortunately, it is difficult to read Excel data into R - Requires additional packages to be installed.
# Convert into CSV and read.


# Built in Data Sets
# Built-in datasets in R
data()# List of built-in Datasets in R

# Loading
data(mtcars)
# Print the first 6 rows
head(mtcars, 6)
```

Descriptive Statistics
```{r}
# Basic Statistics - Mean, Variances,Correlations and T-tests

# Generate a random sample of 100 numbers between 1 and 100
x = sample(x=1:100, size = 100, replace = TRUE)
x # the output of "x" is a vector of data

# Simple Arithmetic Mean
mean(x)

# Calculate Mean when Missing Data is found
y = x # copy x to y
y[sample(x=1:100, size = 20, replace = FALSE)] = NA
y

mean(y)# Will give NA!
# Remove missing value(s)and calculate mean
mean(y, na.rm=TRUE) # Now, it will give the mean value

# Weighted Mean
Grades = c(95,72,87,66)
Weights = c(1/2, 1/4, 1/8, 1/8)
mean(Grades)# Simple Arithmetic mean
weighted.mean(x = Grades, w = Weights)# Weighted Mean

#Variance
var(x)
#Calculating Variance using formula!
sum((x-mean(x))^2)/ (length(x)-1)

# Standard Deviation
sqrt(var(x))
sd(x)
sd(y)
sd(y, na.rm=TRUE)

# Other Commonly Used Functions
min(x)
max(x)
median(x)
min(y)
min(y, na.rm=TRUE)

# Summary Statistics
summary(x)
summary(y)

# Quantiles
quantile(x, probs = c(0.25, 0.75)) # Calculate 25th and 75th Quantile
quantile(x, probs = c(0.1,0.25,0.5, 0.75,0.99))
```

T-test & Annova
```{r}
# T-tests
# Dataset: Tips dependents on...
data(tips, package = "reshape2")
head(tips)
str(tips)

# Gender
unique(tips$sex)
#Day of the week
unique(tips$day)

#One Sample t-test - ONE GROUP [Two Tail. Ho:Mean = 2.5]
t.test(tips$tip, alternative = "two.sided", mu=2.5)
# p value less than Alpha value so we reject the null hypothesis

#One Sample t-test - Upper Tail. Ho:Mean LE 2.5
t.test(tips$tip, alternative = "greater", mu=2.5)
# p value less than Alpha value so we reject the null hypothesis


# Two Sample T-test - TWO GROUP
t.test(tip ~ sex, data = tips, var.equal = TRUE)
# p value Greater than alpha value so we do not reject the null hypothesis i.e. we accept the alternate hypothesis

#ANOVA  - Comparing Multiple Groups
# Tip by the Day of the Week
str(tips)
tipAnova = aov(tip ~ day, tips)
summary(tipAnova)

```

# Learnings from the Assignment
# 1) "#" helps in commenting
# 2) R is Command Line execution i.e. in R every code is executed Line by line as seen in the R markdown
# 3) R is highly case sensitive example: I was not able to execute the "class" function as I capitalised the letter "c" in Class thus you have do be particular. Another Example is in case of Logical Data type i.e. True or False you have to Capitalize the entire word otherwise it will show some error.
# 4) Learnt a lot about how to work with R markdown
# 5) 4 types of Data types - Numeric, Logical, character and date
# 6) 5 types of Data Structure - Vector, Matrix, dataframe, list & array
# 7) There are different packages in R which can be used for different functions as well as for importing different documents.
# 8) You will encounter a lot of problems in your code, you will have to be patient and see every line highlighted in error to correct your code
# 9) You can not learn R or as a matter of fact any programming language in a day. It will take time to be fluent in a R.


